---
title: "tuesday_02_countries"
author: "kelli"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```


## Load Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
#install.packages("ISOcodes")
library(ISOcodes)
library(lubridate)
```



# #triggered

I do not care about countries and their stupid codes in this month of our lord November 2024. 

This is because I want to dig a hole and lie face down until someone in US government  helps the environment or takes money away from creepy and pasty billionaires so that less pasty people can have healthcare.   

Anyway, I am going to put a bunch of quotes from _The English Patient_ by Michael Ondaatje now. Essentially, it's about a guy who got all burned up in Dub Dub Dos and then complained a lot that countries only exist to create wars and hatred. I quote this book I barely remember because the most mature way to process my anger is to be passive-aggressive to the tidy tuesday dataset. 




 <center> <h4> "It's only the rich who can't afford to be smart. They're compromised. They got locked years ago into privilege. They have to protect their belongings. No one is meaner than the rich. Trust me. But they have to follow the rules of their shitty civilised world. They declare war, they have honour, and they can't leave. But you two. We three. We're free." </h4> </center>



I am  going to see how many dead countries there are, and the rough rate of dead countries over the years. I decided to do this because I lived on Johnston Island for six whole months (post death as a country) and that was enough to spur me onward. 





## Get Data

```{r}
former_countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/former_countries.csv')
```


<center> <h4> "Do you understand the sadness of geography?" </center> </h4>


## Clean (suggested code from github)

```{r}
former_countries <-
  ISOcodes::ISO_3166_3 |> 
  tibble::as_tibble() |> 
  janitor::clean_names()
```


## Create a modified dataset to plot
```{r}

types <- former_countries %>% 
  mutate( date = (as.Date(sub("(^[0-9]{4}$)", "\\1-01-01", former_countries$date_withdrawn)))) %>% #Jesus CHRIST that took FOREVER to figure out. This looks for any rows that have four digits only (AKA a single year) and then adds 1-01-01 to it so you can use it as a date.
  select(name, date) %>%
  group_by(date) %>% 
  count( name= "date_count") #count the number of occurances per date
types[,"cum_countries"] <- cumsum(types$date_count) #cumulative sum the occurances
```
 



<center> <h4> “All I desired was to walk upon such an earth that had no maps.”</center> </h4>


## half-hearted graph showing annual country withdrawls and cumulative withdrawls over the dataset.  
```{r suckgraph, fig.path= "../output/"}
ggplot(types)  +  
  geom_bar(aes(x=date, y=date_count, fill = "Yearly Sum"),stat="identity", width = 100)+ 
  geom_line(aes(group = 1, x= date, y = types$cum_countries, color = "Cumulative Sum"), size = 1) +
  labs(title= "Number of countries withdrawn from ISO code standards by year", 
       x="Date",y="Number of Withdrawn Countries") +
  scale_color_manual(values = c("Cumulative Sum" = "black")) +
  labs(color = "", fill = "") +
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_text(size=10),
  axis.text.x = element_text(angle = 45, size=10, vjust = 0.5, hjust=0.5))
  
```


Hi, Kelli here. Thanks for joining me.  I am sorry you have to look upon this snot-fueled pretentious group of quotes from a book I read in 10th grade. It did help a little, though.

## New Things I learned 

1. I used count and cumsum functions! 
2. I made a new Stringr call to fix dates!
3. I made a combo bar AND a line graph! WOW!

