---
title: "An R Markdown to examine the water chemical composition in a Transition zone at differing tide levels"
author: "kelli"
date: "`r Sys.Date()`"
output: 
  html_document:  
    toc: TRUE
    toc_float: TRUE 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../output/")
```



# Setup

### Load libaries 
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(here)
library(kableExtra)
library(ymlthis)
```


### Load **Data**  
```{r dataload}
chemdata <- read_csv(here("week_06", "data", "chemicaldata_maunalua.csv"))
head(chemdata)
```
### Clean this data *up* for a table and a plot. Whee!

Here is how I will do it. 

1. Take out NAs
2. Separate tide and time column
3. For the **table**, I will get the mean of the NN and Phosphate levels in the transition zone by the tide and site.
4. For the **plot**, I will filter so that we only have Nitrogen and Phosphate data from Wailupe.

```{r clean}
### clean data ####
clean <- chemdata %>%
filter(complete.cases(.)) %>% #filters out everything that's not complete
separate(col = Tide_time, #old column
                  into = c("Tide","Time"), #new columns
                  sep = "_")%>%
  select(c(Site, Zone, Tide, Phosphate, NN)) %>%
  filter(Zone == "Transition") #only wanna look at the transition zone
```


# Kable Table
```{r Kable_table}
### Let's make a table of average levels of Phosphate and Nitrate & Nitrite levels by tide and site ###
summary <- clean %>% 
  group_by(Tide, Site) %>%
  summarise(across(where(is.numeric), list(mean = mean), na.rm = TRUE))
kable(summary) %>%
kable_paper("hover", full_width = F)
```
Ok. This table? not great. I am having trouble fixing it up the way I want, like grouping the rows and renaming everything. It is currently 11:46 on Tuesday. I have full confidence I will some day see the light but I might not have time before this is due. I fully accept the consequences of my rushed and botched table. Great job Kel. 

<div class="tenor-gif-embed" data-postid="7414352" data-share-method="host" data-aspect-ratio="0.961538" data-width="100%"><a href="https://tenor.com/view/great-job-good-job-awkward-wink-gif-7414352">Great Job Good Job GIF</a>from <a href="https://tenor.com/search/great+job-gifs">Great Job GIFs</a></div> <script type="text/javascript" async src="https://tenor.com/embed.js"></script>


# PLOT TIME BABY.
We will only plot data from Wailupe. 
```{r plot figure,fig.align='center'}
plot<- clean %>%
  filter(Site == "W")
ggplot(data=plot, mapping= aes(x= log(NN),
                                   y= log(Phosphate),
                                   color = Tide)) + #colors the points by the actual raw data of that column,
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title="Phosphate, Nitrate and Nitrite levels trend higher at low tide",
       x= "Log of Nitrate and Nitrite level",
       y= "Log of Phosphate level",
       color = "Tide",
       caption = "Source: N Silbiger et al., 2018") +
  theme_light()
``` 




# *Wow.* 

We've done some... some work here today. I will understand if I get points knocked off of this one because both that table and that graph contribute nothing new and beautiful to the world we live in. However, it is but a step on the trail to coding transcendence.  
