---
title: "tuesday_04_burders"
author: "kelli"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

## libraries
```{r}
library(tidyverse)
library(tidyplots)
```


The new package I learned this week is tidyplots! Thank you for the last minute shout-out, Dr. Silbiger. 

## read in data
```{r}
episode_metrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-19/episode_metrics.csv')

```

```{r plot,out.width="400%",fig.path = "../output/" }
episode_metrics %>%
  tidyplot(x = episode, y = season, color = exclamation_ratio) %>%
  add_heatmap() +
    scale_x_continuous(breaks = c(1:23),) +
  scale_y_continuous(breaks = c(1:14)) +
  labs(title = "Characters get quieter in Bob's Burgers", caption ="Bob's Burgers goes to a whimper from a bang, according to exclamation point use per episode. Season 8, episode 7 is a Christmas Special lumped with Season 8, episode 6. We were all wondering.", guide_legend("exclamation point ratio per episode"))
```

  scale_x_continuous(breaks = c(1:23),) +
  scale_y_continuous(breaks = c(1:14))
  
 pivot_longer(cols = dialogue_density, unique_words, #columns you want to pivot
               names_to = "Variables", #names of new columns with all the variables
               values_to = "Values") %>% #name of the new column with all the values)
