---
title: "tuesday_05_ICE"
author: "kelli"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Libraries 
```{r}
library(tidyverse)
#install.packages("gifski")
library(gifski)
library(gganimate)
```


## Read in data 
```{r}
cbp_resp <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-26/cbp_resp.csv')
cbp_state <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-26/cbp_state.csv')
```

## First Plot:
I want to see what each region looks like in terms of apprehensions vs. inadmissibles. According to the CPB [here](https://www.cbp.gov/newsroom/stats/cbp-enforcement-statistics/title-8-and-title-42-statistics-fy2021):

> Enforcement Actions refers to apprehensions or inadmissibles processed under CBP’s immigration authority. Inadmissibles refers to individuals encountered at ports of entry who are seeking lawful admission into the United States but are determined to be inadmissible, individuals presenting themselves to seek humanitarian protection under our laws, and individuals who withdraw an application for admission and return to their countries of origin within a short timeframe. Apprehensions refers to the physical control or temporary detainment of a person who is not lawfully in the U.S. which may or may not result in an arrest."

The thing I am most interested in is where "Apprehension" efforts occur more often per method of entry (Southwestern Border, Northern Border, or "Other" assuming that is plane/ship). I am not as interested in title 42 because it is temporary, so I am going to filter out those results.

```{r total, fig.width=11,fig.height=8,fig.align ="center", fig.path = "../output/" }
total_regions <- cbp_resp %>%
  filter(title_of_authority == "Title 8") %>%
  select(area_of_responsibility, demographic, land_border_region, fiscal_year, encounter_count, encounter_type) %>%
  group_by(demographic, land_border_region, fiscal_year, encounter_type) %>%
  summarize(total = sum(encounter_count)) %>%
  ggplot(
    aes(
      x = fiscal_year,
      y = total,
      fill = encounter_type)
    ) +
  geom_col(
    lwd = 0.25, #keep the column widths fairly thin
  ) +
  theme(
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=14), #change legend title font size
        legend.text = element_text(size=10)) + #change legend text font size 
  labs(title = "Where in the US are migrating people physically apprehended?",
       x = "fiscal year",
       y = "number of people CBP 'encountered'") +
  guides(fill=guide_legend(title='Encounter Type')) +
  facet_wrap(~land_border_region)
total_regions
```

Boy... seems like a lot more physical force is being used on the southwestern border, then? Just hypothesizing, but I think that could either be a function of more people crossing illegally, or it could be a function of how much enforcement exists to catch people in each region.


I'm interested to see if all the border offices have increased apprehension numbers as the years go on, or if that trend is different based on office in charge. 

```{r}
categories <- unique(cbp_resp$area_of_responsibility) 
categories
```
Ugh, that's quite a lot of offices. 

Let's just focus on the Southwest, since they have, by far, the most apprehensions.


# Second Plot

I'd like to see if more single adults are apprehended, since I've heard tell from people on TV that people STREAMING over the border are all the "BAD ONES" and not families. I will make a little graphic showing columns of apprehended folks by office, and then a breakdown of their demographic status (Families (FMUA), Single adults, Unaccompanied Children, or Accompanied Children). 


```{r}
filtered_data <- cbp_resp %>%
  filter(title_of_authority == "Title 8",
         land_border_region == "Southwest Land Border") %>%
  select(area_of_responsibility, demographic, citizenship, fiscal_year, encounter_count, encounter_type) %>%
  group_by(demographic, area_of_responsibility, fiscal_year, encounter_type) %>%
  summarize(total = sum(encounter_count))
```

```{r plot, fig.width=11,fig.height=8,fig.align ="center", animation.hook='gifski', interval= 4, fig.path = "../output/"}

plot <- filtered_data %>%
  ggplot(
    aes(
      x = fiscal_year,
      y = total,
      fill = encounter_type)
    ) +
  geom_col(
    lwd = 0.25,
  ) +
  labs(title = "Types of Encounters per Southwest Area of Responsibility",
       x = "fiscal year",
       y = "number of people CBP 'encountered'") +
  guides(fill=guide_legend(title='Encounter Type')) +
  theme(
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=14), #change legend title font size
        legend.text = element_text(size=10)) + #change legend text font size 
  facet_wrap(~area_of_responsibility)
plot

plot2 <- filtered_data %>%
  ggplot(
    aes(
      x = fiscal_year,
      y = total,
      fill = demographic)
    ) +
  geom_col(
    lwd = 0.25,
  ) +
    labs(title = "Demographics Encountered per Southwest Area of Responsibility",
       x = "fiscal year",
       y = "number of people CBP 'encountered'") +
  guides(fill=guide_legend(title='Demographic')) +
  theme(
        legend.key.height = unit(.75, 'cm'), #change legend key height
        legend.key.width = unit(.75, 'cm'), #change legend key width
        legend.title = element_text(size=12), #change legend title font size
        legend.text = element_text(size=8)) + #change legend text font size 
  facet_wrap(~area_of_responsibility)
plot2
```

I suppose my take away from this is that the rise in apprehensions is a trend, but it is not a steadfast one. Also, that it looks like Areas of Responsibility have different duties when it comes to "apprehending" folks or just deeming them inadmissible.

I think it would be interesting to look A) where exactly these areas of responsibility function and B) look at the funding allocated to each office for apprehension efforts. 


# New things I learned: 
1. new package "gifski" makes it easier to use animations in Rmarkdown-- learned how to animate via chunk options 
2. learned how to resize the legends in a plot 
3. used the unique function to list out unique variables in a column 
