---
title: "GoodPlot_BadPlot"
author: "kelli"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


#Load Libraries
```{r}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(ggmap)
library(lubridate)
library(geosphere)
library(ggpubr)
```




#Get Data 
```{r}
meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv")
head(meteorites)

bigfoot <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-13/bigfoot.csv')

```

#Clean up Meteorites
```{r}
cleanm <- meteorites %>%
  filter(lat < 41.01805,
       lat > 36.98442) %>%
  filter(long < -102.0438,
         long > -109.0625) %>%
  select(year, lat, long) %>%
  mutate(sighting = "Meteorite") %>%
  rename(latitude = lat, longitude= long) %>%
  unite(col = "Geopoint", #NEW column
        c(longitude,latitude), #existing columns to unite
        sep = ",", #separating symbol 
        remove = FALSE) #DO NOT delete the original columns
```


# Clean up Bigfoot
```{r}
bigfoot$year <- year(ymd(bigfoot$date)) #extract the year using lubridate
cleanb <- bigfoot %>%
  filter(state == "Colorado")%>%
  select(latitude,longitude,year) %>%
  drop_na(latitude) %>%
  mutate(sighting = "Bigfoot") %>%
  unite(col = "Geopoint", #NEW column
        c(longitude,latitude), #existing columns to unite
        sep = ",", #separating symbol 
        remove = FALSE) #DO NOT delete the original columns

```

#Join Datasets
```{r}
points <- full_join(cleanb,cleanm)
points %>%
  mutate(points, ID = row_number())
```


#Bad Plot 
```{r}
ggplot(data=points,
               aes(x= longitude,
                   y= latitude, color = sighting)) +
  geom_rug() +

  footpic <- image_read("https://www.fortmorgantimes.com/wp-content/uploads/2023/12/thinkstockphotos-152533219.jpg?w=1575") 
```
```



# Good Plot 
```{r}
bbox <- c(left = -109.2, bottom = 36.7, right = -101.9, top = 41.1)

site_map <- ggmap(get_stadiamap(bbox, maptype = "stamen_terrain_background", zoom = 7))+ 
  geom_point(data = cleanb, aes(x = longitude, y = latitude), 
               size = 1.5, color = "orange")+
              geom_point(data = cleanb, aes(x = longitude, y = latitude), 
               pch= 21, size = 2, color = "black")+
    geom_point(data = cleanm, aes(x = longitude, y = latitude), 
               size = 1.5, color = "blue")+
              geom_point(data = cleanm, aes(x = longitude, y = latitude), 
               pch= 21, size = 2, color = "black")+
              theme_bw() +
             labs(title="Bigfeet Avoid Meteorites!!!!!",
       subtitle="I will accept my Nobel prize through the mail, thank you",
       x= "Longitude",
       y= "Latitude")
#ggsave(here("week_07" , "output", "COterrain.png"), width = 11, height = 11) #I think I have screwed up rendering this into a smaller file? I am not sure how to do that with stadia maps yet. 
site_map
```

