---
title: Kelli's Notes for OCN 683
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

###3A: GGPLOT ################
Kelli Konicek 
created on 2024-09-10

#libraries
```{r}
library(tidyverse)
library(here)
library(palmerpenguins)
library(ggplot2)
library(beyonce)
library(ggthemes)
```



LOOK AT YOUR DATA
```{r}
view(penguins)
head(penguins)
tail(penguins)
glimpse(penguins) #ways to take a look at your dataset 
```



#make a plot

```{r}
ggplot(data=penguins) #I know what dataset I'm using but I'm not telling you what to do 
```

make aesthetics- anything that determines how the thing looks based on the data set

```{r}
ggplot(data=penguins, mapping= aes(x= bill_depth_mm,
y= bill_length_mm,
color = species, #colors the points by the actual raw data of that column
shape = island,
size = body_mass_g,
alpha = flipper_length_mm)) +  #alpha changes the density of points
  geom_point() + # plus adds a new function, comma add
  labs(title="Bill depth and length",
       subtitle="Big Bills and Deep Bills for cold ducks",
       x= "Bills Deep (mm)",
       y= "Bills Big (mm)",
       color = "Names of Cold Ducks",
       caption = "Source: Palmer station or citation here") + scale_color_viridis_d()
```

#mapping v setting 
settings are things that are not based ont the values. put settings in the geom_point () layer


```{r}
ggplot(data=penguins, mapping= aes(x= bill_depth_mm,
                                   y= bill_length_mm,
                                   color = species, #colors the points by the actual raw data of that column,
                                   shape = island,
                                   size = body_mass_g,
                                   alpha = flipper_length_mm)) +  #alpha changes the density of points
  geom_point(size=2) + # plus adds a new function, comma add
  labs(title="Bill depth and length",
       subtitle="Big Bills and Deep Bills for cold ducks",
       x= "Bills Deep (mm)",
       y= "Bills Big (mm)",
       color = "Names of Cold Ducks",
       caption = "Source: Palmer station or citation here") + scale_color_viridis_d()
```

#faceting
useful for exploring conditional relationships in large data sets 

###FACET GRID

```{r}
ggplot(data=penguins, mapping= aes(x= bill_depth_mm,
                                   y= bill_length_mm))+
         geom_point() +
         facet_grid(species~sex) #EACH ROW AS A FUNCTION OF COLUMNS
```


###FACET WRAP

#get rid of a legend 
#give more columns to facet
```{r}
ggplot(data=penguins, mapping= aes(x= bill_depth_mm,
                                   y= bill_length_mm))+
  geom_point() +
  facet_wrap(~species) 

# 1D ribbon: YOU determine the dimensions with facet wrap- tilda can make it organized by COLUMNS like this or species~ to make ROWS

ggplot(data=penguins, mapping= aes(x= bill_depth_mm,
                                   y= bill_length_mm,
                                   color= species,
                                   guides(color= FALSE)))+ #GET RID OF THAT LEGEND
  geom_point() +
  facet_wrap(~species, ncol=2) #gives more control with the wrap! 
```


###3B:MORE GGPLOT#######

#geom smooth for regressions
```{r}
ggplot(data=penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm))+
  geom_point()+
  geom_smooth(method="lm", se= FALSE)+ #best way to plot regressions! Whahoo! You can remove the standard error with se=FALSE
  labs(x = "Bill depth (mm)",
       y = "Bill length (mm)")
```






#GROUP BY A COLUMN IN AES
```{r}
ggplot(data=penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     group = species, #GROUP BY SPECIES OR ANOTHER COLUMN WHOO HOO!
                     color = species))+
  geom_point()+
  geom_smooth(method="lm")+ 
  labs(x = "Bill depth (mm)",
       y = "Bill length (mm)")

```


#SCALES

```{r}
ggplot(data=penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     group = species, 
                     color = species))+
  geom_point()+
  geom_smooth(method="lm")+ 
  labs(x = "Bill depth (mm)",
       y = "Bill length (mm)")+
  scale_x_continuous(limits = c(12,20))+ #make a limit for the scales! 
  scale_y_continuous(limits = c(30, 50))
```




#BREAKS and relabling breaks 
```{r}

ggplot(data=penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     group = species, 
                     color = species))+
  geom_point()+
  geom_smooth(method="lm")+ 
  labs(x = "Bill depth (mm)",
       y = "Bill length (mm)")+
  scale_x_continuous(breaks = c(14,17,21),
                     labels = c("low", "medium", "high")) 
  scale_y_continuous(limits = c(30, 50))

```


#MANUALLY CHANGE COLOR
```{r}
  
  ggplot(data=penguins,
         mapping = aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       group = species, 
                       color = species))+
    geom_point()+
    geom_smooth(method="lm")+ 
    labs(x = "Bill depth (mm)",
         y = "Bill length (mm)")+
  scale_x_continuous(breaks = c(14,17,21),
                       labels = c("low", "medium", "high"))+ 
  scale_y_continuous(limits = c(30, 50))+
  scale_color_manual(values = c("orange","blue","black"))
```


#COLOR PALLETS from github

instead of using install.packages, use install_github(USERNAME,PACKAGE NAME)


```{r}
ggplot(data=penguins,
         mapping = aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       group = species, 
                       color = species))+
    geom_point()+
    geom_smooth(method="lm")+ 
    labs(x = "Bill depth (mm)",
         y = "Bill length (mm)")+
    scale_x_continuous(breaks = c(14,17,21),
                       labels = c("low", "medium", "high"))+ 
    scale_y_continuous(limits = c(30, 50))+
    scale_color_manual(values = beyonce_palette(2)) #now you can use the pallette in the place of manual colors!
```




#COORDINATES
#default coordinates for ggplot are cartesian (x,y)
#coord_flip() 
```{r}
  ggplot(data=penguins,
         mapping = aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       group = species, 
                       color = species))+
    geom_point()+
    geom_smooth(method="lm")+ 
    labs(x = "Bill depth (mm)",
         y = "Bill length (mm)")+
    scale_x_continuous(breaks = c(14,17,21),
                       labels = c("low", "medium", "high"))+ 
    scale_y_continuous(limits = c(30, 50))+
    scale_color_manual(values = c("orange","blue","black"))+
    coord_flip()
```


  
#coord_fixed() 
flip the aspect ratio 
```{r}
  ggplot(data=penguins,
         mapping = aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       group = species, 
                       color = species))+
    geom_point()+
    geom_smooth(method="lm")+ 
    labs(x = "Bill depth (mm)",
         y = "Bill length (mm)")+
    scale_x_continuous(breaks = c(14,17,21),
                       labels = c("low", "medium", "high"))+ 
    scale_y_continuous(limits = c(30, 50))+
    scale_color_manual(values = c("orange","blue","black"))+
    coord_fixed()
```


#transform x and y axis 
want to keep the data in its natural scale, but wanna log the visual itself 
  
#coord_trans
coord_trans(x = "log10", y = "log10") 

make the space log times scale. cool! lots of transformations
  
#make coordinates polar 
  #turn x axis into polar coordinates 
  #coord_polar("x")
  
#THEME
composed of four main elements:non data elements you can control
#element_text
  can set the font size and stuff 
#theme() 
allows you to override all the default
complete themes can also be done like theme_gray
 
#axis.title, panel.background
```{r}
  ggplot(data=penguins,
         mapping = aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       group = species, 
                       color = species))+
    geom_point()+
    geom_smooth(method="lm")+ 
    labs(x = "Bill depth (mm)",
         y = "Bill length (mm)")+
    scale_x_continuous(breaks = c(14,17,21),
                       labels = c("low", "medium", "high"))+ 
    scale_y_continuous(limits = c(30, 50))+
    scale_color_manual(values = c("orange","blue","black"))+
    theme_bw()+
    theme(axis.title = element_text(size = 20,
                                    color = "blue"),
          panel.background = element_rect(fill = "linen")) #fill changes the inside, color just changes the outside
```
 

#SAVE THE PLOT: GGSAVE
```{r}
ggsave(here("week_03" , "outputs", "penguins.png"), width = 7, height = 5)
```

  


#OTHER GGPLOT EXTENSIONS
GGPUBR- ggplot extension that makes it publicaion 




#4A: DPLYR ###############


#FILTER

```{r}
library(palmerpenguins)
library(tidyverse)
library(dplyr)
```

```{r}
filter(.data = penguins, sex == "female") 
```

filter(df, type == "otter" & site == "bay")
if the answer is yes, it keeps it-- look at page 8 of Week 4A
== means exactly equal to
asking very explicit 

#list of logical expressions: 
x<y less than 
x%in% y (give it two options-- is it within-- in group membership)
is.na(x) only keep NAs
!is.na(x) do NOT keep NAs
x>y greater than
== exactly equal
x <= y less than or equal to 
x >= y greater than or equal to 
x != y not equal to 


Choose penguins measured in 2008
```{r}
glimpse (penguins)
filter(.data = penguins, year == "2008")
```



Choose penguins with body mass greater than 5000
```{r}
filter(.data = penguins, body_mass_g > 5000)
```


#when do I use quotes? 
If you want to find an exact value or have a character, use quotes. Numbers don't need quotes. Treat things you want EXACTLY like a character.

you can add in multiple layers of things separating by commas 
```{r}
filter(.data = penguins, sex == "female", body_mass_g > 5000)
```


###BOOLEAN OPERATORS

a & b is a AND b
a | b is a OR b 
!a is NOT A 

default for filter is &, but you can use comma for it 

```{r}
filter(.data = penguins, year == "2008" | year == "2009")
```


in group membership (%in%), not(!)
```{r}
filter(.data = penguins, year %in% c("2008", "2009"))

filter(.data = penguins, island != "Dream")

filter(.data = penguins, species %in% c("Adelie", "Gentoo"))
```

penguins between 3000 and 5000 g
```{r}
filter(.data = penguins,
       body_mass_g <5000,
       body_mass_g > 3000)
```




### MUTATE

add a new column converting body mass in g to kg and save the new dataframe

```{r}
data2<-mutate(.data= penguins,
              body_mass_kg = body_mass_g/1000,
              bill_length_depth = bill_length_mm/bill_depth_mm)
```



#add multiple new columns 

mutate multiple columns at the same time

mutate_if(), mutate_at()-- the columns that I tell you, do the same to them, this is good if you have a bunch of columns if you want to do the same thing to them 

mutate_all-- examples on the 4a slide 

#IFELSE()

if this is true, give it whatever value you want, if it is false, give whatever value you want

add new columns based of ddata you already have 

#case_when
ask it for multiple ifelse stuffs -- ifelse is for two things


```{r}
data2 <- mutate(.data = penguins,
                after_2008 = ifelse(year > 2008, "After 2008", "Before 2008"))

data2<-mutate(.data= penguins,
              flipper_plus_body = flipper_length_mm + body_mass_g +
                BIGBOYS = ifelse(body_mass_g > 4000, "BIG BOY", "SMALL BOY")) 
```

      

### THE PIPE

says "and then do" verb 1 and do verb 2 

|>this is also called a NATIVE pipe  magritter pipe needs tidyverse

using the pipe means you don't need the .data 

data3 <- penguins %>%
  filter(sex=="female") %>%
  mutate(log_mass = log(body_mass_g))

### SELECT

#select columns
```{r}
data4 <- penguins %>%
  filter (sex == "female") %>%
  mutate(log_mass = log(body_mass_g) %>%
           select(species, island, sex, log_mass))
```


#RENAME
can use select to RENAME columns, can do it within the select function 
NEW FIRST, THEN OLD 


#SUMMARIZE
have less rows because you are summarizing by a variable you're interested in

want to calculate mean for flipper length 
```{r}
data4 <- penguins %>% 
  summarize(mean_flipper = mean(flipper_length_mm, na.rm = TRUE))
```


you can add mulitple different things to this 
can do summarizeat, summarizeif, summarizeall 

#GROUPBY
almost always associated with summarize 


```{r}
data5 <- penguins %>% 
  group_by(island) %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))

data5 <- penguins %>% 
  group_by(island, sex) %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))

```


#REMOVE NAS

only drop missing data from the column you want

```{r}
penguins %>% 
  drop_na(sex) %>%
  group_by(island, sex) %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))
```



#PIPE TO A GGPLOT
you can clean up all your data and pipe it straight ot GGplot 

YOU GOTTA SWITCH FROM THE PIPE TO THE + WHEN YOU USE GGPLOT


### 4B: TIDYR ###############

#working with tidyr homework 4B: Silbiger et al. biochem data from Hawaii 
#created by kelli 
# last edit 9/24/2024 

#load libaries

```{r}
library(tidyverse)
library(here)
library(ggplot2)
```


#Load data

```{r}
chemdata <- read_csv(here("week_04", "data", "chemicaldata_maunalua.csv"))
view(chemdata)
```


# Clean data: filter, separate
```{r}
clean <- chemdata %>%
  filter(complete.cases(.)) %>% #filters out everything that's not complete
  separate(col = Tide_time, #old column
           into = c("Tide","Time"), #new columns
           sep = "_") #separate old column by this symbol 
view(clean) #looks good 
```
can set "remove = FALSE" to keep the separate column 

#UNITE : brings two columns together 

```{r}
clean <- chemdata %>%
  filter(complete.cases(.)) %>% #filters out everything that's not complete
  separate(col = Tide_time, #old column
           into = c("Tide","Time"), #new columns
           sep = "_") %>% #separate old column by this symbol
  unite(col = "Site_Zone", #NEW column
        c(Site,Zone), #existing columns to unite
        sep = ".", #separating symbol 
        remove = FALSE) #DO NOT delete the original columns 
```


#PIVOT DATA: WIDE TO LONG 
easier to summarize with group_by() and facet_wrap()-- can do these as a function of the variable

#pivot_longer()

```{r}
long <- clean %>%
  pivot_longer(cols = Temp_in:percent_sgd, #columns you want to pivot
               names_to = "Variables", #names of new columns with all the variables
               values_to = "Values") #name of the new column with all the values
long %>%
  group_by(Variables, Site) %>% #group by everything we want 
  summarise(Param_means = mean(Values, na.rm = TRUE), #get mean
            Param_vars = var(Values, na.rm = TRUE))
view(long)
```


THE CHALLENGE 
```{r}
summary <- long %>%
  group_by(Variables, Site_Zone, Tide) %>%
  summarise(Param_means= mean(Values, na.rm = TRUE), #get mean
            Param_vars = var(Values, na.rm = TRUE),
            Param_sd = sd(Values, na.rm = TRUE))
```


#facet_wrap with long data
fix the y axis to let scale be free

```{r}
long %>%
  ggplot(aes(x = Site, y = Values))+
  geom_boxplot()+
  facet_wrap(~Variables, scales = "free")
```

if you want to scale the x, scales = "free_x" or "free_y" to do y only 

#pivot_wider

instead of names_to, we go names_from 

```{r}
wide <- summary %>%
  pivot_wider(names_from = Variables, #column with names for new column 
              values_from = Param_means:Param_sd) #column with the values
```


#WRITE A CSV FILE

```{r}
write_csv("week_04", "output","data","summary.csv")
```



#### 5A: JOINS##################
Learning joins (part of the {dplyr} package)

left_join() -- with real data
Learn to make a tibble

Back to joins (with made up data as examples)

right_join()
inner_join()
full_join()
semi_join()
anti_join()



```{r}
### load libaries ####
library(tidyverse)
library(here)
library(ggplot2)


```
### Load data
```{r}

sites <- read_csv(here("week_05", "data", "site.characteristics.data.csv"))
view(sites)
tpc <-read_csv(here("week_05", "data", "Topt_data.csv"))
glimpse(tpc)
```

site letter links these two data sets together. 

tpc is in wide format, sites is in long 


```{r}
sites_wide <- sites %>%
  pivot_wider(names_from = parameter.measured,
              values_from = values)
view(sites_wide)
```

#put values in alphabetical order by chosen column
but to put it in order of site letter 
```{r}
sites_wide <- sites %>%
  pivot_wider(names_from = parameter.measured,
              values_from = values) %>%
  arrange(site.letter) #arrange data frame by site
view(sites_wide)
```



if you donʻt have any missing values, then all these joins will do the same thing. 


###LEFT JOIN
```{r}
full_left <- left_join(tpc, sites_wide)
view(full_left)
```

left_join(DATA FRAME ON LEFT, DATA FRAME ON RIGHT)
only retains IDS on the left--- will delete 4 if you use left join 
order really matters with the joins. 

###RELOCATE
#look at certain things first or together

helps to clean data, not going to affect analysis 


```{r}
full_left <- full_left <- left_join(tpc, sites_wide) %>%
  relocate(where(is.numeric), .after = where(is.character))
view(full_left)
```



#CALCULATE MEANS AND VARIANCES OF ALL THE NUMERIC DATA 

My first try: 
```{r}
long <- full_left %>%
  pivot_longer(cols = E:substrate.cover, #columns you want to pivot
               names_to = "Numeric_Variables", #names of new columns with all the variables
               values_to = "Values")  #name of the new column with all the values
long %>%
  group_by(Numeric_Variables, site.letter) %>% #group by everything we want 
  summarise(Param_means = mean(Values, na.rm = TRUE), #get mean
            Param_vars = var(Values, na.rm = TRUE))
```

#SUMMARISE AT
```{r}
full_sum <- full_left %>%
  group_by(site.letter)
  summarize_at(vars(E:Tpot, light:substrate.cover),
               funs(mean = mean, var = var), na.rm = TRUE)
```

#SUMMARISE ACROSS
ANOTHER WAY-- find all the numerics, create a list. mean = mean means that you can call your function mean 
```{r}
full_mean_var <- full_left %>%
  group_by(site.letter) %>%
  summarise(across(where(is.numeric), list(mean = mean, var = var), na.rm = TRUE))
```

#SUMMARISE IF
```{r}
summary <- full_left %>%
  select(-site.block)%>%
  group_by(site.letter)%>%
  summarise_if(is.numeric,
               funs(mean=mean, var = var),
               na.rm = TRUE)
```


### DIFFERENT JOINS

creating our own tribble 
make 1 tibble: 

```{r}
T1 <- tibble(Site.ID = c("A", "B", "C", "D"),
Temperature = c(14.1, 15.7, 69, 69.1))

T2 <- tibble(Site.ID = c("A", "B", "C", "D"),
pH = c(14, 1, 2, 5))


```

#INNER JOIN
inner join--- if youʻre going to remove nas anyway you can use this -- will reatain ph data

#FULL JOIN
full join--- keeps literally everything, puts NAs in there
(same as merge but itʻs the tidy version)

#SEMI JOIN-- only going to keep data from first data frame
keeps all the rows from the first data set but ONLY where there is matching data in the second one--- find all the sites that expect in T2, but I only want to keep the temperature data. 

If I donʻt actually want to join everything-- I donʻt want stuff that doesnʻt have pH data 

#ANTI JOIN (T1, T2)
saves all the rows in the first data set that donʻt matching anything in the second data set. 

ORDER MATTERS 



###5B: LUBRIDATE #############
Lubridate date and time

what time is it now? 
```{r}
library(lubridate)
now()
now(tzone = "EST")
now(tzone = "GMT")
```

just the date: 
```{r}
today(tzone = "HST")
am(now()) #is it morning? helpful to divide data 
```

Date specifications for lubridate 
#DATES MUST BE CHARACTER FOR IT TO WORK
i.e. put quotes around it 

#convert a character string into a date: 
```{r}
ymd("2021/02/24")
```

###GUESSING DATES AND TIME FUNCTIONS-- SINGLE DATES
convert to proper ISO date time
year month day hour minute seconds 
```{r}
ymd_hms("2021-02-24 10:22:20 PM")
mdy_hms("02/24/2021 22:22:20")
mdy_hm("February 24 2021 10:22 PM")

#you have to tell it that every piece is there. i.e. if there are seconds, make sure you include there are seconds 
```


### Extracting specific date or time elements from datetimes

#extract the month from the character string 

```{r}
#make a character string of dates
datetimes <- c("02/24/2021 22:22:20",
               "02/25/2021 11:21:10",
               "02/26/2021 8:01:52")

datetimes <- mdy_hms(datetimes)
month(datetimes, label= TRUE, abb = FALSE) #label makes month a FACTOR, abb makes it spell out month by full so that you can make nice clean plots by spelling out the month 
```

#extract the day or the weekday 

```{r}
day(datetimes)
```

#extract the day of the week
```{r}
wday(datetimes, label = TRUE, abbr = FALSE)
```

#extract specific time elements - hour not HourS
```{r}
hour(datetimes)
minute(datetimes)
second(datetimes)
```

#change date and times -- HourS not hour
```{r}
datetimes + hours(4)
datetimes + days(2)
datetimes + minutes(30)
```
#Round Date times 

```{r}
round_date(datetimes, "minute") #round to nearest minute
```

#round to nearest five (however many) minutes
```{r}
round_date(datetimes, "5 mins") #round to nearest FIVE minutes
```

### 6A: R Markdown Intro 
combo of text, code, and output 
whatʻs inside? 
metadata (own language)
text (in markdown)
code (r or python)
output
#Metadata: Yet Another Markup Language (YAML)
---
key: value
---

knitting: share things with collaborators


you can decide whether or not to use the code in the ressults. 

include= FALSE prevents the code AND results frome running 

echo = FALSE: shows results NOT code

message = FALSE: prevents messages generated

warning = FALSE: won't show warnings with code

fig.cap = "..." adds a caption to graphical results 
#6B: RMARKDOWN 2

make comments when you knit that will not be executed 

<!-- This line is commented, won't show in document -->

#Working with Figures
in R markdown, you don't have to use ggsave- can automatically export and fix a figure using Rmarkdown 

change the size in the r command chunk line.  
here are some options: 


i.e. {r, out.width="70%"}

out.width="70%" makes it .7 of width 
fig.width=3 (makes width 3 in)
fig.height=4 (makes height 4 in)
fig.asp.1.2 (changes aspect ratio to 1.2)
fig.align ="center" 

#automatically export figures into same output folder
let's say you want to use one less line of code and just want them all exported to output file no matter what: 

R chunks need to be named for this to work. 
one figure per R chunk. 
set the file path for your fig using: 

fig.path(you can set this within the global settings)

add a label, put it right next to the r in curly brackets 

R automatically saves everything into the FILE directory. but HERE, it's going to use wherever you saved the markdown file as the reference point. 

#automatically save all your figures to your output folder  

cannot use the HERE package like we do with gg save have to put relative paths 



fig.path = "../output/" 

--- "../" 
this sign means go up a directory-- so it's saying "go up from your scripts folder (where this markdown doc is) and go to the output folder to save figures

you can do ../../ to keep going up directories

#save your figure options in markdown 

set a figure caption:

fig.cap ="this is a caption for the image"
if you just want to plot and not the code, set echo=FALSE

save figure as a PDF instead of a PNG
dev="pdf"

#TABLES 
type out tables by hand 

|time | session | topic|
|:---| :--------|------:|
|_left_

Okay this sucks. 

tables with kable 

install.packages("kabelExtra")

#Make a KABLE TABLE 

```{r kabletable}
penguins %>%
  group_by(species) %>%
  summarise(bill_mean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() %>%
  kable_classic_2()%>%
  row_spec(2, bold = T, color = "white",background = "red") %>%
  kable_styling(full_width = FALSE)
```

gt tables is also a really good package and you can add in plots physically in the table 

formattable

dt is a SEARCHABLE TABLE OMG 

can do a cool table as an indepentend project-- redo Karl's table as a searchable table? 

#OUTPUTS 

tons of outputs available too. you can make a github_document you can knit and view the html on githubt without pasting into the raw.githack package

some of these options might not be an option depending on the output, you have to look at the YAML arguments 

#help with YAML 

ymlthis is the package
can create yaml for different output packages 

clock "toosl and then ADDINS then scroll to "write new RMarkdown on YAML file" 




Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

