---
title: "week_13A_homework"
output:
  github_document

# Libraries

```{r}
library(tidyverse)
library(here)

```

## Instructions

You have a set of 4 .csv files in data/homework. Each of these files is
a timeseries of temperature and light data collected in tide pools in
Oregon by Jenn Fields.

Your goal is to bring in all 4 files and calculate the mean and standard
deviation of both temperature (Temp.C) and light (Intensity.lux) for
each tide pool.

Use both a for loop and map() functions in your script. (Basically, do
it twice).

# For Loop

## Using loops to read in multiple .csv files

### Read in one of the files so that you can see what it looks like.

```{r}
testdata<-read_csv(here("week_13", "data", "homework","TP1.csv"))
glimpse(testdata)
```

### list files in a directory

go to this folder and bring in all the data in this folder

```{r}
# point to the location on the computer of the folder
homePath<-here("Week_13", "data", "homework")
# list all the files in that path with a specific pattern

# In this case we are looking for everything that has a .csv in the filename

# you can use regex to be more specific if you are looking for certain patterns in filenames

files <- dir(path = homePath,pattern = ".csv")
files
```

### pre-allocate space for the loop

```{r}
# pre-allocate space
# make an empty dataframe that has one row for each file and  columns
 home_data<-tibble(filename =  rep(NA, length(files)),  # column name for file
                   mean_temp = rep(NA, length(files)),
                   sd_temp = rep(NA, length(files)),# column name for the mean temperature
                   mean_lux = rep(NA, length(files)),
                   sd_lux = rep(NA, length(files))# column name for the mean salinity
                   ) # column name for the year name
home_data
```

### test

write basic code to calculate a mean and build out

```{r}
raw_data<-read_csv(paste0(homePath,"/",files[1])) 
```

great, we can go to the next step

```{r}
mean_temp<-mean(raw_data$Temp.C, na.rm = TRUE) # calculate a mean
mean_temp
```

### Turn it into a for loop

```{r}
for (i in 1:length(files)){ # loop over 1:4 the number of files
}

```

the only thing that is being iterated here is the file, not saving the
raw data

Add in the loop over the raw data

```{r}
for (i in 1:length(files)){ # loop over 1:4 the number of files 
raw_data<-read_csv(paste0(homePath,"/",files[i]))}
```

## Add in means and standard devs to the loop

```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(homePath,"/",files[i]))
#glimpse(raw_data)
home_data$filename[i]<-files[i]
home_data$mean_temp[i]<-mean(raw_data$Temp.C, na.rm =TRUE)
home_data$sd_temp[i]<- sd(raw_data$Temp.C, na.rm = TRUE)
home_data$mean_lux[i]<-mean(raw_data$Intensity.lux, na.rm =TRUE)
home_data$sd_lux[i]<- sd(raw_data$Intensity.lux, na.rm = TRUE)
} 
home_data

```

# Purrr

## Bring in files using purrr instead of a for loop

### find the files

```{r}
# point to the location on the computer of the folder
housePath<-here("week_13", "data", "homework")
files <- dir(path = housePath,pattern = ".csv")
files
```

gives you entirety of the path for each one

```{r}
files <- dir(path = housePath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files
```

### read in the files


```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename
data
```

### calculate means 

```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temp.C, na.rm = TRUE),
            sd_temp = sd(Temp.C, na.rm = TRUE),
            mean_lux = mean(Intensity.lux, na.rm = TRUE),
            sd_lux = sd(Intensity.lux ,na.rm = TRUE))
data
```

# YEEEE HAW!
