---
title: "tuesday_03_monster"
author: "kelli"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


# Load Libraries
```{r}
library(tidyverse)
library(plotly)
library(here)
```


# Get Data
```{r getit}
monster_movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movies.csv')
```


# Plot that data- with point labels! 
```{r plot}
p <- ggplot(data = monster_movies, aes(x = average_rating, y = num_votes, text =      paste("primary_title:", primary_title))) + #this is how you can make point text
     geom_point(alpha = (1/3)) + scale_x_log10() + #scale points and decrease alpha
  labs(title = "What Monster-ish movies are truly popular?",
       x = "Average Rating",
       y = "Number of Votes")

ggplotly(p)

ggsave(here("tidy_tuesday" , "tuesday_03", "output", "monster.png"), width = 7, height = 5) #This is a dumb .png but juuuust in case I need an output! 
```

I have learned that Monsterʻs Inc. is truly the most popular of the Monster-inclusive Movies, because it has the highest rating with the most votes by far. I am not surprised. 

However, I am more interested in movies that got a lot of bad ratings, because I like Bad Movies. Robot Monster (3.0 rating, 5538 votes), Monsters in the Woods (2.6 rating, 1497 votes), and Bottle Monster (2.5 rating at 932 votes) are where my heart lies. 


# What I learned 
I have never used the Plotly package! Being able to zoom into figures and see text above a point is a game changer. I love this thing and I am glad I tried it out.


