---
title: "07A_Homework: Make a map using tidy tuesday data"
author: "kelli"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../output/")
```

# Set up
## Load Libraries
```{r libraries}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(ggmap)

```




## Get Data 
```{r data}
bigfoot <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-13/bigfoot.csv')

```

# I'd like to take you on a short journey.

I grew up in Colorado, and I wanted to know how often we get Bigfoot sightings as a state. 

I found something that **blew the lid** off of Bigfoot in Colorado. 

## Filter bigfoot data to Colorado sightings only 
```{r}
clean <- bigfoot %>%
  filter(state == "Colorado")
```

## Map Colorado Bigfoot sightings onto a polygon. 
```{r COpolygon, fig.cap= "You see this? DO YOU SEE THIS? The sightings *stop* as you got East."}
states <- map_data("state")

CO <- states %>%
  filter(region == "colorado")
ggplot()+
  geom_polygon(data = CO, 
               aes(x = long, 
                   y = lat, 
                   group = group),
               color = "black",
               fill = "beige")+
  theme_minimal()+
  guides(fill = FALSE)+
  theme(panel.background = element_rect(fill = "brown"))+
    coord_map(projection = "mercator",
            xlim = c(-102.0438,-109.0625),
            ylim = c(37, 41)) +
  geom_point(data = clean, # add a point at all my sites
             aes(x = longitude,
                 y = latitude))+
  labs(title="Bigfoot sightings in Colorado",
       subtitle="SHOCKING trend revealed",
       x= "Latitude",
       y= "Longitutde")
```
These results are *shocking*... unless you know something else about Colorado.  



### Colorado? It has mountains. And Plains. 



I decided to investigate further to see if most bigfoot sightings occur in the mountains, as opposed to the plains area of Colorado. And WOW. What I found will ASTOUND YOU. 


# UNDISSPUTED Proof of Big feet Populations

I decided to use a terrain map that shows, roughly, where the Rocky Mountains begin in Colorado. For the layperson who isn't an R master and expert cartographer like myself, Green on map means mountain zone. 


```{r COterrain, fig.cap= "Look at it! LOOK AT IT!"}
bbox <- c(left = -109.0625, bottom = 37, right = -101, top = 41)

site_map <- ggmap(get_stadiamap(bbox, maptype = "stamen_terrain_background", zoom = 7))+ #I got this from internet and it worked! Whee! 
  geom_point(data = clean, aes(x = longitude, y = latitude), 
               size = 1.5, color = "orange")+
              geom_point(data = clean, aes(x = longitude, y = latitude), 
               pch= 21, size = 2, color = "black")+ #outline helps show individual points
              theme_bw() +
             labs(title="Adding terrain to map of Colorado shows Bigfoot is real",
       subtitle="I will accept my Nobel prize through the mail, thank you",
       x= "Latitude",
       y= "Longitutde")
#ggsave(here("week_07" , "output", "COterrain.png"), width = 11, height = 11) #I think I have screwed up rendering this into a smaller file? I am not sure how to do that with stadia maps yet. 
site_map
```


As you can see, The Bigfeeet reside mainly in the mountainous regions, probably _because it is more difficult to spot them._ They are *so smart* and *so real* that they know where the best habitat is to remain undetected. 

### It is truly amazing how data can show us The Truth.